---
title: "p8105_hw1_zt2397"
output: github_document
date: "2026-09-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

## Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```

```{r}
str (penguins)
levels (penguins$species)
levels (penguins$island)
```
The dataset 'penguins' contains data from three islands, "Biscoe," "Dream," and "Torgersen." It also contains three penguin species "Adelie," "Chainstrap," and "Gentoo."
The dataset contains 'r nrow(penguins)' and 'r ncol(penguins)'
Important variables include bill length, bill depth, flipper length, and body mass in grams, as well as sex. 
The mean bill length is 'r mean(penguins$flipper_length_mm, na.rm = TRUE)

### Problem 1 Scatterplot
```{r}
flipper_bill_scatterplot=
ggplot(penguins,
      aes(x = bill_length_mm,
        y = flipper_length_mm,
        color = species))+ geom_point() +
  labs ( # Used this to give labels to the scatterplot
    title = "Flipper Length vs Bill Length", 
    x = "Bill Length (mm)",
    y = "Flipper Length (mm)"
    )

flipper_bill_scatterplot
```
```{r}
ggsave("flipper_bill_scatterplot.png",
       plot = flipper_bill_scatterplot)
```

## Problem 2

```{r}
random_sample = rnorm (10)

random_sample
positive = random_sample > 0 
positive

character_vector = c("apple", "banana", "carrot", "spinach", "salmonn", "quinoa", "almonds", "yogurt", "broccoli", "lentils")
character_vector

food_group = factor (c ("fruit", "fruit", "vegetable", "vegetable", "protein", "protein", "protein", "protein", "vegetable","protein"))
food_group
```
```{r}
problem2_p2 = 
  tibble (
    random_sample = random_sample,
    positive = positive,
    character_vector = character_vector,
    food_group = food_group
  )
problem2_p2
```
```{r}
mean(pull(problem2_p2, random_sample))
mean(pull(problem2_p2, positive))
mean(pull(problem2_p2, character_vector))
mean(pull(problem2_p2, food_group))
```
Taking the mean of "random sample" works as expected since it is numeric data, and taking the mean of "positive also works since it treats categorical data as 0 or 1. However, character vector has no numeric meaning as it is text, and for "food groups" it is stored internally.
```{r}
as.numeric (positive)
as.numeric (character_vector)
as.numeric (food_group)
```

Applying "as.numeric" converts all of the values to numeric values, such as the fruits, vegetables, and protein are assigned 1,2, and 3. This helps explain why mean fails on charcater vector as these values do not represent any meaningful numbers, so averaging creates non-meaningful results. 

