---
title: "p8105_hw1_zt2397"
output: github_document
date: "2026-09-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

## Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```

```{r}
str (penguins)
levels (penguins$species)
levels (penguins$island)
```
The dataset 'penguins' contains data from three islands, "Biscoe," "Dream," and "Torgersen." It also contains three penguin species "Adelie," "Chainstrap," and "Gentoo."
The dataset contains 'r nrow(penguins)' and 'r ncol(penguins)'
Important variables include bill length, bill depth, flipper length, and body mass in grams, as well as sex. 
The mean bill length is 'r mean(penguins$flipper_length_mm, na.rm = TRUE)

### Problem 1 Scatterplot
```{r}
flipper_bill_scatterplot=
ggplot(penguins,
      aes(x = bill_length_mm,
        y = flipper_length_mm,
        color = species))+ geom_point() +
  labs ( # Used this to give labels to the scatterplot
    title = "Flipper Length vs Bill Length", 
    x = "Bill Length (mm)",
    y = "Flipper Length (mm)"
    )

flipper_bill_scatterplot
```
```{r}
ggsave("flipper_bill_scatterplot.png",
       plot = flipper_bill_scatterplot)
```



